### 2023-02-11 presentation/slides/slides.Rmd
updating tables

---
title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India"
subtitle: 
author: 
  - "Joshua D. Merfeld <br> KDI School and IZA"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(icons)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')
# Double check
getwd()    # check

reportP <- function(pValue){
  if (pValue < 0.001){
    result <- "p < 0.001"
  } else {
    result <- sprintf("p = %.3f", pValue) # inserts a float into a string and simultaneously do rounding
  }
  return(result)
}

reportDec <- function(dec){

    result <- sprintf("%.3f", dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

reportDecOne <- function(dec){

    result <- sprintf("%.1f", dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

reportPercent <- function(dec){

    result <- sprintf("%.1f", 100*dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

commaNum <- function(large){

  result <- formatC(large, big.mark = ",")

  return(result)
}

# Shapefile of india
india_shape <- read_sf("data/spatial/shapefiles/state.shp")

# Coal plants
### Load raw data ----------------------------------------------------------------------------------

plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
# India only
plants <- plants %>% filter(Country=="India")
# Also want only things with a non-missing year
plants <- plants %>% filter(is.na(Year)==F)

# Just keep what we want 
plants <- plants %>% dplyr::select(plant_id = ParentID, 
                                   unit_id = Tracker.ID, 
                                   capacity = `Capacity.(MW)`, 
                                   year_built = Year, 
                                   year_retired = RETIRED, 
                                   lat = Latitude, 
                                   lon = Longitude)

plants <- plants %>% mutate(capacity = capacity/1000)

# long
plantsyearly <- c()
for (year in 1990:2015){
  plantstemp <- plants %>% filter(year_built<=year & (year_retired>year | is.na(year_retired)==TRUE))
  plantstemp <- plantstemp %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup() %>%
                  dplyr::select(plant_id, capacity, lat, lon)
  plantstemp$year <- year
  plantsyearly <- rbind(plantsyearly, plantstemp)
}

# Wind example
# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
# Go through two months worth of days
date <- as_date("2010-01-01")
villageswind <- c()
for (add_day in 1:100){
  wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-", month(date), "-", day(date), ".csv")) %>% as_tibble()
  wind <- wind[,1:2]
  colnames(wind) <- c("shrid", "indicator")
  villagestemp <- villages %>% left_join(wind, by = "shrid")
  villagestemp$date <- date

  villageswind <- rbind(villageswind, villagestemp)

  date <- date + 1
}

villageswind <- villageswind %>% st_set_crs(st_crs(districts))

# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
# style_duo_accent(
#   primary_color = "#006334",
#   secondary_color = "#A7A9AC",
#   inverse_header_color = "#FFFFFF"
# )
# style_mono_light(
#                  base_color = "#006334"
#                  )
style_duo(
          primary_color = "#A7A9AC", 
          secondary_color = "#006334"
          )
```

```{css, echo=F}
    /* Table width = 100% max-width */

    <!-- .remark-slide table{ -->
    <!--     width: 100%; -->
    <!-- } -->

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    tfoot {
      font-size: 80%;
    }
```

## Coal and energy worldwide

- Coal one of most common energy sources worldwide
  - Reached an all-time high in 2021 <small><small>(www.iea.org/reports/coal-fired-electricity)</small></small>

![]("../../../pollution_development/draft/tables/energy.png")

---

## Coal and energy in India

![]("../../../pollution_development/draft/tables/energy.png")

---

## Coal and energy in India

- From 1990 to 2010: coal generation from 42.4 gigawatts to more than 100
- Obvious benefits to more power <small><small>(Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017)</small></small>

---

## Downsides, too

- But power generation can have downsides, too
  - Especially true near plants: pollution a major concern
- Climate change and pollutants like SO2, mercury, etc. <small><small>(www.eia.gov/energyexplained/coal/coal-and-the-environment.php)</small></small>
- Long literature on negative effects of pollution 
  - Health <small><small>(Arceo, Hanna, and Oliva 2016; Brunekreef and Holgate 2002; Deryugina et al. 2019; Kampa and Castanas 2008; Pope III and Dockery 2006)</small></small>
  - Productivity/labor allocation <small><small>(Chang et al. 2019; Ebenstein, Lavy, and Roth 2016; Hanna and Oliva 2015; He, Liu, and Salvo 2019; Wen and Burke 2022)</small></small>

---

## This paper

- Effects of pollution from coal plants on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - One additional day of exposure per season: decrease of 0.3 percent (yields)

---

## This paper

- Adults decrease labor allocation
  - Driven by large decrease in older half of sample; younger half increases allocation to farm labor
- More pollution in previous year also leads to lower nightlight growth
  - 25th to 75th percentile of exposure leads to decrease in nightlights of 0.7 percent

---

## Data sources

```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

data <- matrix(NA, nrow = 8, ncol = 3)
rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
colnames(data) <- c("source", "geographic coverage", "temporal coverage")
data[1,1] <- "Asher et al. (2021)"
data[2,1] <- "Global Energy Monitor"
data[3,1] <- "NCAR"
data[4,1] <- "Hammer et al. (2020)"
data[5,1] <- "Angopadhyay et al. (2022)"
data[6,1] <- "Asher et al. (2021)"
data[7,1] <- "TerraClimate"
data[8,1] <- "National Sample Survey (NSS)"
data[1,2] <- "India"
data[2,2] <- "global"
data[3,2] <- "global"
data[4,2] <- "global"
data[5,2] <- "India"
data[6,2] <- "India"
data[7,2] <- "global"
data[8,2] <- "India"
data[1,3] <- " "
data[2,3] <- "yearly"
data[3,3] <- "daily"
data[4,3] <- "monthly"
data[5,3] <- "two seasons/year"
data[6,3] <- "yearly"
data[7,3] <- "monthly"
data[8,3] <- "daily interviews"

# Table
kable(
      data,
      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
      ) %>%
  footnote(
           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2,width = "4.5cm") %>%
  column_spec(c(3:4),width = "3.5cm") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_classic_2()

```

---

## General idea
- Coal plants open in a given year
- Plot wind direction each day
  - Exposure means wind blowing from coal plant to village (within 30km)
  - Aggregate up to desired level (e.g. season, month, week)

---

## Coal plants over time

```{r plants, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = india_shape, color = "gray85", fill = "gray80") +
    geom_point(data = plantsyearly, 
               aes(x = lon, y = lat, size = capacity),
               colour = "#006334", alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = year,
                      cumulative = FALSE
                      ) +
    labs(title = "year: {current_frame}") +
    theme(legend.position = "none")

```

---

## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r windbase, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = districts, color = "gray85", fill = "gray80") +
    geom_sf(data = villageswind %>% filter(date=="2010-01-01"),
            alpha = .5) +
    labs(title = "date: 2010-01-01") +
    theme(legend.position = "none")

```

---

## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r wind, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = villageswind, 
            aes(fill = indicator),
            alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = date,
                      cumulative = FALSE
                      ) +
    labs(title = "date: {current_frame}") +
    theme(legend.position = "none")

```

---

## Where do coal plants open?

```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable,
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  footnote(
           "Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE,
           escape = FALSE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

```

---

## Validating wind direction - particulate matter and exposure

```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable,
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

```

---

## Effects of wind direction on agricultural productivity

```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtable,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(11, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

---

## Effects of PM on agricultural productivity

- Likely endogenous
  - Upward biased?

```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtablepm[-c(3:6),],
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(3, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

---

## Effects of PM on agricultural productivity

- We are really interested in the effect of _pollution_, not wind
- Use wind as instrument? Assumes...
  - Conditional on fixed effects and weather...
  - Wind only affects agriculture through pollution
- Reasonable?

---

## Effects of PM on agricultural productivity, IV estimates

```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtabletwo[-c(3:6, 12, 16:19),],
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(c(3,8), bold = TRUE) %>%
  row_spec(c(2), hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

---

## Robustness checks

- Leads of wind direction
- Only post coal plant opening

---

## Taking stock

- We see clear negative effects of pollution on productivity
  - Absolute deviation of exposure is approx. 8 days, meaning 2-3 percent change in productivity
  - Using AD of PM, it's above 20 percent change
- Key question: what is the driver?
  - Land?
  - Labor?
- Hard to look explicitly at land, so let's look at labor

---

## National Sample Survey (NSS)

```{r nss, echo = FALSE, message = FALSE, warning = FALSE,}

laborplot

```

---

## Problem: only have district identifiers, so aggregate up

```{r aggregation, echo = FALSE, message = FALSE, warning = FALSE,}

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')

# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
wind <- wind %>% dplyr::select(shrid, days_sum)
villages <- villages %>% left_join(wind, by = "shrid")
villages <- villages %>% st_set_crs(st_crs(districts))

ggplot() + 
  theme_map() +
  geom_sf(data = districts, fill = NA) +
  geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
  scale_fill_viridis_c("exposure (days)", option = "viridis") +
  labs(x = "lon",
       y = "lat") +
  theme(legend.position = c(0.5, 0.05),
        legend.direction = "horizontal",
        legend.key.width = unit(0.3, 'cm'),
        legend.text = element_text(size = 8),
        legend.background=element_blank() )

```

---

## Wind direction and labor allocation, by age

```{r labortableold, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
labortableold <- labortableold[,-1]
colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
kable(
      labortableold,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:6),width = "1.5cm") %>%
  row_spec(c(10), hline_after = TRUE) %>%
  row_spec(c(6,9), bold = TRUE) %>%
  kable_classic_2()

```

---

## Large negative effects on older half

- Older half of sample decreases overall labor allocation
  - Driven by self-employment and nonfarm employment
- Younger half actually increases allocation!
- Consistent with health effects only on older adults, though not conclusive

---

## Effects on aggregate growth?

- We know pollution affects productivity and labor allocation
- Might it also affect growth?
- Proxy for growth with nightlights
  - Look at previous year's exposure and current year's nightlights
  - "Growth" of nightlights
  - Village FE

---

## Wind direction and growth of nightlights

```{r ntltable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(ntltable) <- c("(1)", "(2)")
kable(
      ntltable,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(9), hline_after = TRUE) %>%
  row_spec(c(5,8), bold = TRUE) %>%
  kable_classic_2()

```

---

## Conclusion

- Large growth in coal power in India
  - Other countries, as well
- Increase in electricity, but downsides as well
  - Pollution!
- We already have plenty of evidence on effects of pollution on health

---

## Conclusion

- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
  - Identification relies on within-village changes across seasons
- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
- Clear changes in labor allocation
  - Oldest half of sample
- Slower growth of nightlights

---

## Conclusion

- Overall cost/benefit way beyond scope of paper
- Instead, results point to importance of the location of pollution sources
  - Downwind villages suffer
- 30km radius indicates relatively long-range effects