### 2023-02-21 presentation/slides/slides.Rmd
lots of updates!

---
title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India"
subtitle: 
author: 
  - "Joshua D. Merfeld <br> KDI School and IZA"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(icons)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')
# Double check
getwd()    # check

reportP <- function(pValue){
  if (pValue < 0.001){
    result <- "p < 0.001"
  } else {
    result <- sprintf("p = %.3f", pValue) # inserts a float into a string and simultaneously do rounding
  }
  return(result)
}

reportDec <- function(dec){

    result <- sprintf("%.3f", dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

reportDecOne <- function(dec){

    result <- sprintf("%.1f", dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

reportPercent <- function(dec){

    result <- sprintf("%.1f", 100*dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

commaNum <- function(large){

  result <- formatC(large, big.mark = ",")

  return(result)
}

# Shapefile of india
india_shape <- read_sf("data/spatial/shapefiles/state.shp")

# Coal plants
### Load raw data ----------------------------------------------------------------------------------

plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
# India only
plants <- plants %>% filter(Country=="India")
# Also want only things with a non-missing year
plants <- plants %>% filter(is.na(Year)==F)

# Just keep what we want 
plants <- plants %>% dplyr::select(plant_id = ParentID, 
                                   unit_id = Tracker.ID, 
                                   capacity = `Capacity.(MW)`, 
                                   year_built = Year, 
                                   year_retired = RETIRED, 
                                   lat = Latitude, 
                                   lon = Longitude)

plants <- plants %>% mutate(capacity = capacity/1000)

# long
plantsyearly <- c()
for (year in 1990:2015){
  plantstemp <- plants %>% filter(year_built<=year & (year_retired>year | is.na(year_retired)==TRUE))
  plantstemp <- plantstemp %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup() %>%
                  dplyr::select(plant_id, capacity, lat, lon)
  plantstemp$year <- year
  plantsyearly <- rbind(plantsyearly, plantstemp)
}

# Wind example
# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
# Go through two months worth of days
date <- as_date("2010-01-01")
villageswind <- c()
for (add_day in 1:100){
  wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-", month(date), "-", day(date), ".csv")) %>% as_tibble()
  wind <- wind[,1:2]
  colnames(wind) <- c("shrid", "indicator")
  villagestemp <- villages %>% left_join(wind, by = "shrid")
  villagestemp$date <- date

  villageswind <- rbind(villageswind, villagestemp)

  date <- date + 1
}

villageswind <- villageswind %>% st_set_crs(st_crs(districts))

# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
labortablerural <- readRDS("pollution_development/draft/tables/labortablerural.rds")
labortableurban <- readRDS("pollution_development/draft/tables/labortableurban.rds")
labortableruralyoung <- readRDS("pollution_development/draft/tables/labortableruralyoung.rds")
labortableruralold <- readRDS("pollution_development/draft/tables/labortableruralold.rds")
wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
labortableoldstate <- readRDS("pollution_development/draft/tables/labortableoldstate.rds")
ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
yieldtablemonthly <- readRDS("pollution_development/draft/tables/yieldtablemonthly.rds")

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
# style_duo_accent(
#   primary_color = "#006334",
#   secondary_color = "#A7A9AC",
#   inverse_header_color = "#FFFFFF"
# )
# style_mono_light(
#                  base_color = "#006334"
#                  )
style_duo(
          primary_color = "#A7A9AC", 
          secondary_color = "#006334"
          )
```

```{css, echo=F}
    /* Table width = 100% max-width */

    <!-- .remark-slide table{ -->
    <!--     width: 100%; -->
    <!-- } -->

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    tfoot {
      font-size: 80%;
    }
```

## Coal and energy worldwide

- Coal one of most common energy sources worldwide
  - Reached an all-time high in 2021 <small><small>(www.iea.org/reports/coal-fired-electricity)</small></small>

![]("../../../pollution_development/draft/tables/energy.png")

---

## Coal and energy in India

![]("../../../pollution_development/draft/tables/energy.png")

---

## Coal and energy in India

- From 1990 to 2010: coal generation from 42.4 gigawatts to more than 100
- Obvious benefits to more power <small><small>(Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017)</small></small>

---

## Downsides, too

- But power generation can have downsides, too
  - Especially true near plants: pollution a major concern
- Climate change and pollutants like SO2, mercury, etc. <small><small>(www.eia.gov/energyexplained/coal/coal-and-the-environment.php)</small></small>
- Long literature on negative effects of pollution 
  - Health <small><small>(Arceo, Hanna, and Oliva 2016; Brunekreef and Holgate 2002; Deryugina et al. 2019; Kampa and Castanas 2008; Pope III and Dockery 2006)</small></small>
  - Productivity/labor allocation <small><small>(Chang et al. 2019; Ebenstein, Lavy, and Roth 2016; Hanna and Oliva 2015; He, Liu, and Salvo 2019; Wen and Burke 2022)</small></small>

---

## This paper

- Effects of pollution from coal plants on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - One additional day of exposure per season: decrease of 0.3 percent (yields)

---

## This paper

- Adults decrease labor allocation
  - Driven by large decrease in older half of sample; younger half increases allocation to farm labor
- More pollution in previous year also leads to lower nightlight growth
  - 25th to 75th percentile of exposure leads to decrease in nightlights of 0.7 percent

---

## Data sources

```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

data <- matrix(NA, nrow = 8, ncol = 3)
rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
colnames(data) <- c("source", "geographic coverage", "temporal coverage")
data[1,1] <- "Asher et al. (2021)"
data[2,1] <- "Global Energy Monitor"
data[3,1] <- "NCAR"
data[4,1] <- "Hammer et al. (2020)"
data[5,1] <- "Angopadhyay et al. (2022)"
data[6,1] <- "Asher et al. (2021)"
data[7,1] <- "TerraClimate"
data[8,1] <- "National Sample Survey (NSS)"
data[1,2] <- "India"
data[2,2] <- "global"
data[3,2] <- "global"
data[4,2] <- "global"
data[5,2] <- "India"
data[6,2] <- "India"
data[7,2] <- "global"
data[8,2] <- "India"
data[1,3] <- " "
data[2,3] <- "yearly"
data[3,3] <- "daily"
data[4,3] <- "monthly"
data[5,3] <- "two seasons/year"
data[6,3] <- "yearly"
data[7,3] <- "monthly"
data[8,3] <- "daily interviews"

# Table
kable(
      data,
      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
      ) %>%
  footnote(
           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2,width = "4.5cm") %>%
  column_spec(c(3:4),width = "3.5cm") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_classic_2()

```

---

## General idea
- Coal plants open in a given year
- Plot wind direction each day
  - Exposure means wind blowing from coal plant to village (within 30km)
  - Aggregate up to desired level (e.g. season, month, week)

---

## Coal plants over time

```{r plants, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = india_shape, color = "gray85", fill = "gray80") +
    geom_point(data = plantsyearly, 
               aes(x = lon, y = lat, size = capacity),
               colour = "#006334", alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = year,
                      cumulative = FALSE
                      ) +
    labs(title = "year: {current_frame}") +
    theme(legend.position = "none")

```

---

## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r windbase, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = districts, color = "gray85", fill = "gray80") +
    geom_sf(data = villageswind %>% filter(date=="2010-01-01"),
            alpha = .5) +
    labs(title = "date: 2010-01-01") +
    theme(legend.position = "none")

```

---

## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r wind, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = villageswind, 
            aes(fill = indicator),
            alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = date,
                      cumulative = FALSE
                      ) +
    labs(title = "date: {current_frame}") +
    theme(legend.position = "none")

```

---

## Where do coal plants open?

```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable,
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  footnote(
           "Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE,
           escape = FALSE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

```

---

## Validating wind direction - particulate matter and exposure, monthly level

```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable,
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

```

---

## Effects of wind direction on agricultural productivity

```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtable,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(11, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

---

## Effects of PM on agricultural productivity

- Likely endogenous
  - Upward biased?

```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtablepm[-c(3:6),],
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(3, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

---

## Effects of PM on agricultural productivity

- We are really interested in the effect of _pollution_, not wind
- Use wind as instrument? Assumes...
  - Conditional on fixed effects and weather...
  - Wind only affects agriculture through pollution
- Reasonable?

---

## Effects of PM on agricultural productivity, IV estimates

```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtabletwo[-c(3:6, 12, 16:19),],
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(c(3,8), bold = TRUE) %>%
  row_spec(c(2), hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

---

## Robustness checks

- Leads of wind direction
- Only post coal plant opening

---

## Agricultural productivity and pollution leads/lags

```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtableleads) <- c("wind", "IV - PM (log)")
yieldtableleads[1:6,2] <- yieldtableleads[7:12,2]
yieldtableleads <- yieldtableleads[-c(7:12),]
yieldtableleads <- yieldtableleads[-c(7:10),]
rownames(yieldtableleads)[c(1,3,5)] <- c("lag", "current", "lead")
kable(
      yieldtableleads,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  kable_classic_2()

```

---

## Taking stock

- We see clear negative effects of pollution on productivity
  - Absolute deviation of exposure is approx. 8 days, meaning 2-3 percent change in productivity
  - Using AD of PM, it's above 20 percent change
- Key question: what is the driver?
  - Land?
  - Labor?
- Hard to look explicitly at land, so let's look at labor

---

## National Sample Survey (NSS)

```{r nss, echo = FALSE, message = FALSE, warning = FALSE,}

laborplot

```

---

## Problem: only have district identifiers, so aggregate up

```{r aggregation, echo = FALSE, message = FALSE, warning = FALSE,}

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')

# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
wind <- wind %>% dplyr::select(shrid, days_sum)
villages <- villages %>% left_join(wind, by = "shrid")
villages <- villages %>% st_set_crs(st_crs(districts))

ggplot() + 
  theme_map() +
  geom_sf(data = districts, fill = NA) +
  geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
  scale_fill_viridis_c("exposure (days)", option = "viridis") +
  labs(x = "lon",
       y = "lat") +
  theme(legend.position = c(0.5, 0.05),
        legend.direction = "horizontal",
        legend.key.width = unit(0.3, 'cm'),
        legend.text = element_text(size = 8),
        legend.background=element_blank() )

```

---

## Wind direction and labor allocation, by rural status

```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labortablerural) <- c("all", "all", "self", "wage", "farm", "non-farm")
colnames(labortableurban) <- c("all", "all", "self", "wage", "farm", "non-farm")

mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat1) <- c("rural")
mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat2) <- c("urban")

labortablerural2 <- rbind(mat1, labortablerural[c(1:2, 7:8),], mat2, labortableurban[-c(3:6),], labortableurban[c(3:6),])

kable(
      labortablerural2[-c(nrow(labortablerural2)-2, nrow(labortablerural2)-1, nrow(labortablerural2)),],
      align = "cccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. The outcome in each regression is days worked in the previous seven days. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:7),width = "1.5cm") %>%
  kable_classic_2()

```

---

## Wind direction and labor allocation, by age (rural only)

```{r labortableruralage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labortableruralyoung) <- c("all", "all", "self", "wage", "farm", "non-farm")
colnames(labortableruralold) <- c("all", "all", "self", "wage", "farm", "non-farm")

mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat1) <- c("younger than 35")
mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat2) <- c("45 or older")

labortableruralyoung2 <- rbind(mat1, labortableruralyoung[c(1:2, 7:8),], mat2, labortableruralold[-c(3:6),], labortableruralold[c(3:6),])

kable(
      labortableruralyoung2[-c(nrow(labortableruralyoung2)-2, nrow(labortableruralyoung2)-1, nrow(labortableruralyoung2)),],
      align = "cccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. The outcome in each regression is days worked in the previous seven days. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:7),width = "1.5cm") %>%
  kable_classic_2()

```

---

## One last check

- Farm labor most common in July 
- Have monthly values of wind and pollution...
  - Heterogeneity by month of the season?
  - Monsoon season only

---

## Heterogeneity by month of the monsoon season

```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtablemonthly) <- c("(1)", "(2)")

kable(
      yieldtablemonthly[c(1:10, nrow(yieldtablemonthly)),],
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "1.5cm") %>%
  kable_classic_2()

```

---

## Effects on aggregate growth?

- We know pollution affects productivity and labor allocation
- Might it also affect growth?
- Proxy for growth with nightlights
  - Look at previous year's exposure and current year's nightlights
  - "Growth" of nightlights
  - Village FE

---

## Wind direction and growth of nightlights

```{r ntltable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
ntltable <- ntltable[,c(1,3)]
colnames(ntltable) <- c("(1)", "(2)")
kable(
      ntltable,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(9), hline_after = TRUE) %>%
  row_spec(c(5,8), bold = TRUE) %>%
  kable_classic_2()

```

---

## Conclusion

- Large growth in coal power in India
  - Other countries, as well
- Increase in electricity, but downsides as well
  - Pollution!
- We already have plenty of evidence on effects of pollution on health

---

## Conclusion

- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
  - Identification relies on within-village changes across seasons
- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
- Clear changes in labor allocation
  - Oldest half of sample
- Slower growth of nightlights

---

## Conclusion

- Overall cost/benefit way beyond scope of paper
- Instead, results point to importance of the location of pollution sources
  - Downwind villages suffer
- 30km radius indicates relatively long-range effects